Bayesian Shape Reconstruction and Optimal Guidance for Autonomous Landing on Asteroids

### 2021-12-22 manuscript.tex
Update affliation and change objective function J

% todo: we may update the affiliation and the address of Shankar 
\institute{S. Kulumani (corresponding author) \at
              Mechanical and Aerospace Engineering, The George Washington University, Washington DC, 20052\\
              \email{shanks.k@gmail.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           T. Lee,  \at
              Mechanical and Aerospace Engineering, The George Washington University, Washington DC, 20052\\
              \email{tylee@gwu.edu}
}

@@

Using Hamilton's principle one can derive the inertial equations of motion of the dumbbell spacecraft~\cite{kulumani2017b} as
\begin{align}
    \dot{\vc{x}} &= \vc{v}, \label{eq:x_dot}\\
    \parenth{m_1 + m_2} \dot{\vc{v}} &= m_1 R_A \deriv{U}{\vc{z}_1} + m_2 R_A \deriv{U}{\vc{z}_2} + u_f,\label{eq:v_dot} \\
    \dot{R} &= R S(\vc{\Omega}) , \label{eq:R_dot}\\
    J \dot{\vc{\Omega}} + \vc{\Omega} \times J \vc{\Omega} &= \vc{M}_1 + \vc{M}_2 + u_m. \label{eq:omega_dot}
\end{align}
The vectors \( \vc{z}_1 \) and \( \vc{z}_2\in\R^3\) define the position of the dumbbell masses in the asteroid fixed frame and are defined as
\begin{align}
    \vc{z}_1 &= R_A^T \parenth{\vc{x} + R \vc{\rho}_1} , \\
    \vc{z}_2 &= R_A^T \parenth{\vc{x} + R \vc{\rho}_2},
\end{align}
where \( \rho_i \in\R^3 \) defines the position of each mass in the spacecraft fixed body frame.
The control inputs to the spacecraft are defined by \( u_f, u_m \in \R^3 \) which define the control force represented in the inertial frame and the control moment represented in the spacecraft frame, respectively.
The standard moment of inertia of the dumbbell model is 
\begin{align}\label{eq:dumbbell_moment_of_inertia}
    J_I = \sum_i^n J_i + m_i \parenth{\vc{\zeta}_i^T \vc{\zeta}_i I_{3\times 3} - \vc{\zeta}_i \vc{\zeta}_i^T} , 
\end{align}
where \( \vc{\zeta}_i \) is the position of \( m_i \) in the spacecraft fixed frame and the moment of inertia of each sphere is
\begin{align}\label{eq:sphere_moment_of_inertia}
    J_i = \begin{bmatrix} 
        \frac{2}{5} m_i r_i^2 & 0 & 0 \\
        0 & \frac{2}{5} m_i r_i^2 & 0 \\
        0 & 0 & \frac{2}{5} m_i r_i^2 
    \end{bmatrix}.
\end{align}
\Cref{eq:dumbbell_moment_of_inertia} is consistent with the well-known parallel-axis theorem~\cite{greenwood1988}.
We utilize a nonstandard moment of inertia in the equations of motion, \( J_d \in \R^{3 \times 3 } \), which is related to the standard moment of inertia by
\begin{subequations}\label{eq:moi_transformation}
    \begin{align}
        J &= \tr{J_d} I_{3\times 3} - J_d, \label{eq:moi_ns2s}\\
        J_d &= \frac{1}{2} \tr{J} I_{3\times 3} - J.\label{eq:moi_s2ns}
    \end{align}
\end{subequations}     
The gravitational moment on the dumbbell \( \vc{M}_i\) is defined as
\begin{align}
    \vc{M}_i = m_i \parenth{S(R_A^T \vc{\rho}_i) R^T \deriv{U}{\vb{z}_i}}.
\end{align}
In the above equations, the gravitational potential is denoted by $U\in\R$, which is computed by a polyhedron model as described in the subsequent section.

@@

We define a cost associated with each vertex \( \vc{v}_i \) of the shape estimate as
\begin{align}\label{eq:explore_cost}
    J_i (x, R, R_A) = \alpha_w J_{w_i} + \alpha_d J_{d_i}(x_r) + \alpha_c J_{c_i}(x_r)
\end{align}
where the weighting factors \( \alpha_w, \alpha_d, \alpha_c \in \R \) are chosen such that \( \alpha_w + \alpha_d + \alpha_c = 1 \).
The cost function is defined as a function of the current inertial position, \( x \in \R^3 \), and the attitude, \( R \in \SO \) of the spacecraft.
Furthermore, the knowledge of the small body rotation is required in order to determine the position of the spacecraft in the small body fixed frame, \( x_r = R_A^T x\).

The term \( J_{w_i} \in \R^1 \) represents the cost associated with the uncertainty of vertex \( i \) as
\begin{align}\label{eq:weight_cost}
    J_{w_i} &= - \frac{w_i}{w_m}
\end{align}
where \( w_i \) is the uncertainty of vertex \( i \), which is defined as the variance of the radius in the preceding section, and \( w_m \in\R \) is a maximum uncertainty used to scale the values.
The term \( J_{d_i} \) represents the scaled geodesic distance between the current state of the spacecraft and vertex \( i \),
\begin{align}\label{eq:distance_cost}
    J_{d_i}(\rpos) &= \frac{1}{\pi} \arctan \parenth{ \frac{\norm{\rpos \times \vc{v}_i}}{\rpos \cdot \vc{v}_i}}.
\end{align}

@@

We assume $\theta$ varies linearly with respect to $t$, and substitute the desired trajectory into \Cref{eq:v_dot} to obtain the control force required to follow the desired trajectory in the absence of any tracking error:
\begin{align}\label{eq:tracking_control_cost}
    u_f(\theta) = -F_{ext}(x_d(\theta)), 
\end{align}
which is computed by the polyhedron potential model given in~\cref{eq:attraction}.
The control cost is then defined as the integral over the desired trajectory~\cref{eq:spherical_waypoint} between the current state and the desired vertex as
\begin{align}\label{eq:control_cost}
    J_{c_i}(\rpos) = \frac{1}{u_m} \int_{\theta_0}^{\theta_f} u_f(\theta)^T R u_f(\theta) d\theta,
\end{align}
where \( u_m \) is used to normalize and scale \( J_{c_i} \).
\Cref{eq:control_cost} is numerically integrated over the trajectory \( x_d(t) \) and used to penalize vertices which have a larger cost.

The vertex which minimizes~\cref{eq:explore_cost} 
\begin{align*}
    \vc{v}_{min} = \argmin_{i} \{J_i(x, R, R_A)\},
\end{align*}
is determined.
This vertex is then used to determine the desired trajectory of the spacecraft in order to collect a measurement as in~\Cref{eq:spherical_waypoint}.

@@

Utilizing these metrics, a landing site is chosen to minimize the surface cost given as
\begin{align}\label{eq:surface_cost}
    \min_{v_i} J_l(x, v_i) =  J_{\text{distance}}(x, v_i) - J_{\text{science}}(v_i),
\end{align}
subject to a hard inequality constraint requiring that the surface slope is less than a threshold, i.e., \( \phi \leq \phi_m \).
The surface cost described here is relatively simple and chosen to be solely a function of the current spacecraft position and shape model.
Additional scientific criteria such as surface material or location on the body, e.g. equatorial vs. polar regions, can be utilized instead of a random science value as presented here.

>>>>

% todo: we may update the affiliation and the address of Shankar 
\institute{S. Kulumani (corresponding author) \at
              The Johns Hopkins University, Applied Physics Laboratory, Laurel, MD 20723\\
              \email{shankar.kulumani@jhuapl.edu}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           T. Lee,  \at
              Mechanical and Aerospace Engineering, The George Washington University, Washington DC, 20052\\
              \email{tylee@gwu.edu}
}

@@

Using Hamilton's principle one can derive the inertial equations of motion of the dumbbell spacecraft~\cite{kulumani2017b} as
\begin{align}
    \dot{\vc{x}} &= \vc{v}, \label{eq:x_dot}\\
    \parenth{m_1 + m_2} \dot{\vc{v}} &= m_1 R_A \deriv{U}{\vc{z}_1} + m_2 R_A \deriv{U}{\vc{z}_2} + u_f,\label{eq:v_dot} \\
    \dot{R} &= R S(\vc{\Omega}) , \label{eq:R_dot}\\
    J \dot{\vc{\Omega}} + \vc{\Omega} \times J \vc{\Omega} &= \vc{M}_1 + \vc{M}_2 + u_m. \label{eq:omega_dot}
\end{align}
The vectors \( \vc{z}_1 \) and \( \vc{z}_2\in\R^3\) define the position of the dumbbell masses in the asteroid fixed frame and are defined as
\begin{align}
    \vc{z}_1 &= R_A^T \parenth{\vc{x} + R \vc{\rho}_1} , \\
    \vc{z}_2 &= R_A^T \parenth{\vc{x} + R \vc{\rho}_2},
\end{align}
where \( \rho_i \in\R^3 \) defines the position of each mass in the spacecraft fixed body frame.
The control inputs to the spacecraft are defined by \( u_f, u_m \in \R^3 \) which define the control force represented in the inertial frame and the control moment represented in the spacecraft frame, respectively.
The standard moment of inertia of the dumbbell model is 
\begin{align}\label{eq:dumbbell_moment_of_inertia}
    J_I = \sum_i^n J_i + m_i \parenth{\vc{\zeta}_i^T \vc{\zeta}_i I_{3\times 3} - \vc{\zeta}_i \vc{\zeta}_i^T} , 
\end{align}
where \( \vc{\zeta}_i \) is the position of \( m_i \) in the spacecraft fixed frame and the moment of inertia of each sphere is
\begin{align}\label{eq:sphere_moment_of_inertia}
    J_i = \begin{bmatrix} 
        \frac{2}{5} m_i r_i^2 & 0 & 0 \\
        0 & \frac{2}{5} m_i r_i^2 & 0 \\
        0 & 0 & \frac{2}{5} m_i r_i^2 
    \end{bmatrix}.
\end{align}
\Cref{eq:dumbbell_moment_of_inertia} is consistent with the well-known parallel-axis theorem~\cite{greenwood1988}.
% We utilize a nonstandard moment of inertia in the equations of motion, \( J_d \in \R^{3 \times 3 } \), which is related to the standard moment of inertia by
% \begin{subequations}\label{eq:moi_transformation}
%     \begin{align}
%         J &= \tr{J_d} I_{3\times 3} - J_d, \label{eq:moi_ns2s}\\
%         J_d &= \frac{1}{2} \tr{J} I_{3\times 3} - J.\label{eq:moi_s2ns}
%     \end{align}
% \end{subequations}     
The gravitational moment on the dumbbell \( \vc{M}_i\) is defined as
\begin{align}
    \vc{M}_i = m_i \parenth{S(R_A^T \vc{\rho}_i) R^T \deriv{U}{\vb{z}_i}}.
\end{align}
In the above equations, the gravitational potential is denoted by $U\in\R$, which is computed by a polyhedron model as described in the subsequent section.

@@

We define a cost associated with each vertex \( \vc{v}_i \) of the shape estimate as
\begin{align}\label{eq:explore_cost}
    \mathcal{J}_i (x, R, R_A) = \alpha_w \mathcal{J}_{w_i} + \alpha_d \mathcal{J}_{d_i}(x_r) + \alpha_c \mathcal{J}_{c_i}(x_r)
\end{align}
where the weighting factors \( \alpha_w, \alpha_d, \alpha_c \in \R \) are chosen such that \( \alpha_w + \alpha_d + \alpha_c = 1 \).
The cost function is defined as a function of the current inertial position, \( x \in \R^3 \), and the attitude, \( R \in \SO \) of the spacecraft.
Furthermore, the knowledge of the small body rotation is required in order to determine the position of the spacecraft in the small body fixed frame, \( x_r = R_A^T x\).

The term \( \mathcal{J}_{w_i} \in \R^1 \) represents the cost associated with the uncertainty of vertex \( i \) as
\begin{align}\label{eq:weight_cost}
    \mathcal{J}_{w_i} &= - \frac{w_i}{w_m}
\end{align}
where \( w_i \) is the uncertainty of vertex \( i \), which is defined as the variance of the radius in the preceding section, and \( w_m \in\R \) is a maximum uncertainty used to scale the values.
The term \( \mathcal{J}_{d_i} \) represents the scaled geodesic distance between the current state of the spacecraft and vertex \( i \),
\begin{align}\label{eq:distance_cost}
    \mathcal{J}_{d_i}(\rpos) &= \frac{1}{\pi} \arctan \parenth{ \frac{\norm{\rpos \times \vc{v}_i}}{\rpos \cdot \vc{v}_i}}.
\end{align}

@@

We assume $\theta$ varies linearly with respect to $t$, and substitute the desired trajectory into \Cref{eq:v_dot} to obtain the control force required to follow the desired trajectory in the absence of any tracking error:
\begin{align}\label{eq:tracking_control_cost}
    u_f(\theta) = -F_{ext}(x_d(\theta)), 
\end{align}
which is computed by the polyhedron potential model given in~\cref{eq:attraction}.
The control cost is then defined as the integral over the desired trajectory~\cref{eq:spherical_waypoint} between the current state and the desired vertex as
\begin{align}\label{eq:control_cost}
    \mathcal{J}_{c_i}(\rpos) = \frac{1}{u_m} \int_{\theta_0}^{\theta_f} u_f(\theta)^T R u_f(\theta) d\theta,
\end{align}
where \( u_m \) is used to normalize and scale \( J_{c_i} \).
\Cref{eq:control_cost} is numerically integrated over the trajectory \( x_d(t) \) and used to penalize vertices which have a larger cost.

The vertex which minimizes~\cref{eq:explore_cost} 
\begin{align*}
    \vc{v}_{min} = \argmin_{i} \{\mathcal{J}_i(x, R, R_A)\},
\end{align*}
is determined.
This vertex is then used to determine the desired trajectory of the spacecraft in order to collect a measurement as in~\Cref{eq:spherical_waypoint}.

@@

Utilizing these metrics, a landing site is chosen to minimize the surface cost given as
\begin{align}\label{eq:surface_cost}
    \min_{v_i} \mathcal{J}_l(x, v_i) =  \mathcal{J}_{\text{distance}}(x, v_i) - \mathcal{J}_{\text{science}}(v_i),
\end{align}
subject to a hard inequality constraint requiring that the surface slope is less than a threshold, i.e., \( \phi \leq \phi_m \).
The surface cost described here is relatively simple and chosen to be solely a function of the current spacecraft position and shape model.
Additional scientific criteria such as surface material or location on the body, e.g. equatorial vs. polar regions, can be utilized instead of a random science value as presented here.
